For the iCMs data set we will go through similar steps as for the aCMs data to create the QFeatures object.

```{r, eval=FALSE}
# Reading in necessary libraries
library("scp")
library("tidyverse")
```

Reading in the data.

```{r, eval=FALSE}
icmsTab <- read_tsv("iCMs.tsv.gz")
```

Like for the aCMs data, we can create an initial QFeatures object to check how many experiments are in there. However, this is a much larger data set and will therefore take more time to load. We did it and already know that we expect there to be 2184 experiments.

```{r,eval=FALSE}
# icms <- readSCPfromDIANN(icmsTab)
```

For this dataset, we also want to extract metadata from the file names. We want to extract:

-   Sample name

-   Date

-   Day

-   Batch

-   Platewell

-   Position

-   Number of cells analysed

```{r,eval=FALSE}
# Extracting Sample name, Date, Day, Batch, Platewell, Position and number of cells analysed
tab_icms <- tibble(File.Name = unique(icmsTab[[1]]))|>
  mutate(Sample = sub(".*\\\\", "",File.Name))|>
  mutate(Date = sub("^.*?(\\d{6}).*", "\\1", File.Name))|>
  mutate(Date = ymd(as.integer(Date)))|>
  mutate(Batch = sub("^.*?(Batch[0-9])_.+$", "\\1", File.Name))|>
  mutate(PlateWell = sub("^.*?([A-Z][0-9]+)_.+$", "\\1", File.Name))|>
  mutate(Position = sub("^.*?([A-Z]+[0-9]+)_1_[0-9]+\\.d$", "\\1", File.Name))|>
  mutate(Day = sub("^.*?(day[0-9]+)_.+$", "\\1", File.Name, ignore.case = T))|>
  mutate(Day = sub("\\D+", "\\1", Day))|>
  mutate(Day = as.numeric(Day))|>
  mutate(Cell_count = sub("^.*?([0-9]+cell).+$", "\\1", File.Name))
```

Getting an overview of the metadata we collected.

```{r,eval=FALSE}
table(tab_icms$Date)
table(tab_icms$Batch)
table(tab_icms$Day)
table(tab_icms$Cell_count)

table(tab_icms$Date,tab_icms$Batch)
table(tab_icms$Day,tab_icms$Date)
table(tab_icms$Batch,tab_icms$Cell_count)
```

The output should look like this:

```{r,eval=FALSE}
> table(tab_icms$Date)
2023-06-16 2023-06-27 2023-07-11 2023-07-26 2023-08-15 2023-08-21 
       358        377        367        378        323        381 

> table(tab_icms$Batch)
Batch1 Batch2 Batch3 Batch4 Batch5 Batch6 
   377    323    381    367    378    358 

> table(tab_icms$Day)
 day0 day10  day2 day21  day4 
  358   755   381   323   367 

> table(tab_icms$Cell)
10cell  1cell 50cell 
   287   1611    286 
```

```{r,eval=FALSE}
> table(tab_icms$Date,tab_icms$Batch)
             Batch1 Batch2 Batch3 Batch4 Batch5 Batch6
  2023-06-16      0      0      0      0      0    358
  2023-06-27    377      0      0      0      0      0
  2023-07-11      0      0      0    367      0      0
  2023-07-26      0      0      0      0    378      0
  2023-08-15      0    323      0      0      0      0
  2023-08-21      0      0    381      0      0      0
  
> table(tab_icms$Day,tab_icms$Date)
     2023-06-16 2023-06-27 2023-07-11 2023-07-26 2023-08-15 2023-08-21
  0         358          0          0          0          0          0
  2           0          0          0          0          0        381
  4           0          0        367          0          0          0
  10          0        377          0        378          0          0
  21          0          0          0          0        323          0
  
> table(tab_icms$Batch,tab_icms$Cell_count)
         10cell 1cell 50cell
  Batch1     48   285     44
  Batch2     48   226     49
  Batch3     48   283     50
  Batch4     47   273     47
  Batch5     48   282     48
  Batch6     48   262     48

```

Adding the `runCol` column data frame for creating the QFeatures object.

```{r,eval=FALSE}
tab_icms$runCol <- tab_icms$File.Name
```

Creating the `QFeatures` object with the extracted metadata.

```{r,eval=FALSE}
icms <- readSCPfromDIANN(icmsTab,
                         colData = DataFrame(tab_icms))
```

Setting the sample names of the `QFeatures` object.

```{r,eval=FALSE}
names(icms) <- icms$Sample
```

Replace 0 by NA for downstream analysis.

```{r,eval=FALSE}
icms <- zeroIsNA(icms, names(icms))
```

Removing all precursors that have been mapped to protein groups.

```{r,eval=FALSE}
icms <- filterFeatures(icms, ~ !grepl(";", Protein.Names))
```

Replacing row names with the precursor identifications.

```{r,eval=FALSE}
expl <- lapply(experiments(icms),
               function(x) {
                 rownames(x) <- rowData(x)$Precursor.Id
                 return(x)
               })
experiments(icms) <- List(expl)
```

Joining all the assays based on the precursors

```{r,eval=FALSE}
icms <- joinAssays(icms, i = names(icms), name = "precursors")
```
