
```{r}
library(QFeatures)
library(scp)
library(dplyr)
```

```{r}
library(scp)
library(data.table)
library(ggplot2)
library(patchwork)
library(SingleCellExperiment)
library(scater)

## Get PSM data
report <- fread("~/Documents/.localData/SCP/cardiomiocytes-single-cell-proteomics-data/aCMs/report.tsv", sep = "\t")
## Get annotations
annot <- DataFrame(File.Name = unique(report$File.Name))
moreAnnots <- do.call(rbind, strsplit(annot$File.Name, "_"))
annot$Date <- sub("PRO.*\\\\", "", moreAnnots[, 2])
annot$Subject <- sub(".*([1-4]).*", "subject\\1", moreAnnots[, 4])
annot$Location <- tolower(moreAnnots[, 6])
annot$Location[annot$Location %in% c("1", "d")] <- "sytox"
annot$PlateWell <- moreAnnots[, 7]
annot$PlateRow <- gsub("[0-9]*", "", annot$PlateWell)
annot$PlateCol <- gsub("[A-Z]*", "", annot$PlateWell)
annot$Undocument <- moreAnnots[, 8]
annot$FileIndex <- sub(".d", "", moreAnnots[, 10])
annot$runCol <- annot$File.Name

## Format to QFeatures
acms <- readSCPfromDIANN(report, colData = annot)
acms <- zeroIsNA(acms, names(acms))


acms <- readRDS("data/acms.rds")
acms <- acms[, , 1:299]
acms$FileIndex <- as.numeric(sub(".*_(\\d*).d", "\\1", acms$File.Name))
acms$PlateColumn <- as.numeric(sub("[A-Z]", "", acms$PlateWell))
```


# Quality control 

The objective of quality control is to remove low-quality data. These
data can be filtered based on PSM-level annotations and on cell-level
annotations.

## PSM filtering

We first explore the features annotation. We can retrieve all the data
annotations along the quantified values in a single long table using
`longFormat()`

```{r}
lf <- longFormat(
    acms, colvars = c("HeartLocation", "Subject", "FileIndex"),
    rowvars = c("PEP", "RT", "Predicted.RT", "IM", "Predicted.IM")
)
```

We first explore the distributions of the posterior error
probabilities (PEP) for precursor identification for each cell
separately, stratified by subject, a potential variable for batch
effects. We sort the cells according to the sequence of MS acquisition
(available in the `FileIndex` column).

```{r, fig.width=10, fig.height=4}
lf <- data.frame(lf) |> 
    mutate(FileIndex = factor(FileIndex, levels = sort(unique(FileIndex))))
ggplot(lf) +
    aes(y = log10(PEP), 
        x = FileIndex,
        colour = Subject) +
    geom_boxplot()
```

The plot indicates that the identification errors of every precursor
are stable across the experiment. We will therefore not use PEP for
PSM filtering.

We now generate a similar plot, but focusing on the difference between
the observed retention time (RT) and the predicted RT, where large
differences may indicate wrong identification or issues during liquid
chromatography. We hence plot the absolute difference between observed
RT.

```{r, fig.width=8, fig.height=3}
ggplot(lf) +
    aes(y = abs(RT - Predicted.RT), 
        x = FileIndex,
        colour = Subject) +
    geom_boxplot() +
    geom_hline(yintercept = 0.18, linetype = "dashed") +
    scale_y_log10()
```

Again, the plot shows stable metrics across cells, with a few outliers
above 0.18 spread across the experiment. We will remove those outliers
using `filterFeatures()` that takes the `QFeatures` object and an
arbitrary formula to filter the PSM. Note that you can directly filter
on variables from the rowData by including them in the formula.

```{r}
acms <- filterFeatures(acms, ~ abs(RT - Predicted.RT) < 0.18)
```

Next, we remove precursors that map on multiple proteins, as these are
of limited use when interpreting downstream, we will remove them.
DIA-NN generated the `isProteotypic`, however the table below shows
that some precursors considered as proteotypic may still be part of a
protein group where the `Protein.Ids` contain multiple proteins
separated by a `;`.

```{r}
table(isProteotypic = rd$Proteotypic, 
      isProteinGroup = grepl(";", rd$Protein.Ids))
```

We here keep all precursor that are both proteotypic and not part of a
protein group.

```{r}
acms <- filterFeatures(acms, ~ !grepl(";", Protein.Names) &
                           Proteotypic == 1)
```

Finally, we add the information whether a precursor belongs to a
contaminant protein or not. We retrieve this information using the
cRAP database, through the `camprotR` package.

```{r}
library("camprotR") ## "CambridgeCentreForProteomics/camprotR"
contaminants <- get_ccp_crap()
head(contaminants)
```

We next retrieve all the rowData table in a single table using
`rbindRowData()`, we add a new column `isContaminant` and inject it 
back into the `QFeatures` object after splitting it by assay.

```{r}
rd <- rbindRowData(acms, names(acms))
rd$isContaminant <- rd$Protein.Ids %in% contaminants
rowData(acms) <- split(rd, rd$assay)
```

We decide not to remove contaminants as they may be used later for
model exploration and diagnostics.

## Sample filtering

We now remove low-quality cells. We will compute a few metrics,
explore the experimental design and assess how it impacts the computed
metrics and we will conclude with removing cells identified as
low-quality, if any.

### Compute QC metrics

First, we compute the median intensity within each cell. This is meant
as a proxy for the amount of material that has been injected in the
MS. Note that in single-cell experiments, difference in the amount of 
material injected are subject to technical variability, but also to
biological variability (cell size, proteome complexity, cell
density, ...). Hence, different median intensities may not be a
problematic artifact if different across heart location.

```{r}
acms$MedianIntensity <- sapply(names(acms), function(i) {
    median(log2(assay(acms[[i]])), na.rm = TRUE)
})
```

We also compute the number of identified precursors within each cell.
This relates to the number of rows in each set of the `QFeatures`
object.

```{r}
acms$TotalIds <- nrows(acms)
```

### Explore the experimental design

We now explore the experimental design. First, the sample annotation 
contain subjects and different dates. The table below shows that each
subject has been processed in a different day. 

```{r}
table(acms$Date, acms$Subject)
```

This is not a problematic design as soon we are not interested in
distinguishing between subject effects and date effects, which is the
case for this experiment.

Next we will explore the plate design. We first retrieve all the
available sample annotations from the `colData` and perform a little
data cleaning for improved visualisation. 

```{r}
plateData <- data.frame(colData(acms)) |> 
    mutate(PlateCol = factor(acms$PlateWell, levels = 1:24))
```

We will first explore how the data acquisition strategy, by plotting 
the order of the the MS runs on the plate layout.

```{r, fig.width=9, fig.height=3}
ggplot(plateData) +
    aes(y = PlateLetter, 
        x = PlateColumn, 
        fill = FileIndex) +
    geom_tile() +
    facet_grid(~ Date) +
    scale_fill_continuous(type = "viridis") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "bottom")
```

We can see that each plate as been acquired sequentially (as already
suggested by the date annotation), in order of the rows. Next, we now
plot the plate layout with respect to the heart location of the cells,
for each data/subject separately.

```{r, fig.width=9, fig.height=3}
ggplot(plateData) +
    aes(y = PlateRow, 
        x = PlateCol, 
        fill = Location) +
    geom_tile() +
    facet_grid(~ Date) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "bottom")
```

Two observations can be drawn from the plot. First, all `sytox` cells
are from one date and hence their effect cannot be estimated.
According to the original study, this sample preparation for this
subject failed. Hence we will remove the corresponding data later.
Also cells have been assigned in block on the plate, one block for
each heart location. While this is probably the consequence of
practical constraints, this means we can no longer distinguish between
well column effect (if any) and heart location effects. Hence, we will
ignore cell column effects. 

We plot the same layout, this time colouring by the number of
precursors identified.

```{r, fig.width=9, fig.height=3}
ggplot(plateData) +
    aes(y = PlateRow, 
        x = PlateCol, 
        fill = TotalIds) +
    geom_tile() +
    facet_grid(~ Date) +
    scale_fill_continuous(type = "viridis") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "bottom")
```

Columns 10-24 in the first plate seem to have more identified 
precursors than the other columns, potentially indicating some batch
effect during cell sorting. A cell type effect may also be an
explanation, but we do not see the effect in the last plate.

```{r, fig.width=9, fig.height=3}
ggplot(plateData) +
    aes(y = PlateLetter, 
        x = PlateColumn, 
        fill = MedianIntensity) +
    geom_tile() +
    facet_grid(~ Date) +
    scale_fill_continuous(type = "viridis") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "bottom")
```

Interestingly, this plot shows that columns 1-9 in the first plate
have lower median intensity. We now combine the information 

```{r}
ggplot(plateData) +
    aes(y = MedianIntensity,
        x = TotalIds,
        colour = FileIndex,
        shape = Subject) +
    geom_point(size = 2) +
    scale_colour_continuous(type = "viridis")
```

### Filter samples

Upon data exploration, we identified that one of the subjects should
be removed.

```{r}
acms <- subsetByColData(acms, acms$Subject != "Subject5")
```

## Log-transformation

A last step before modelling is to log-transform the data.    

```{r}
acms <- logTransform(acms, "precursors", "precursors_log")
```
